### 2022-12-12 tex/lenv.tex
[ edit ] some corrections and notes after meeting
[ edit ] new draft of chapters 5 and 6

@@

We could, indeed, use these three clauses to define what an environment is.
However, I find them difficult to work with, as it is often easier to do
linear algebraic proofs separately from the rest of an environment.
For identity and composition, as we are about to see, the original definition
is easier to use because we can rely on the identity and composition of linear
maps.
Concretely, an inductive proof of identity would, for example, involve
constructing an environment of type
$\grP\gamma, \grQ\delta \env\V \grP\gamma, \grQ\delta$ by constructing
environments of types $\grP\gamma, \gr0\delta \env\V \grP\gamma$ and
$\gr0\gamma, \grQ\delta \env\V \grQ\delta$.
These are not identity environments, so we would have to strengthen the
induction hypothesis.

One of the primary test cases for environments is simultaneous substitution,
which will look like the following rule.
The admissibility of substitution will be by induction on the derivation of
$\Delta \vdash A$, so we will need to be able to adapt any environment we are
given to work with any possible context of new premises.
In the simply typed case, the only change to the context we encountered was the
binding of new variables.
Now, with usage annotations, we furthermore have linear decompositions of the
context, necessitating changes to the environment whenever usage annotations
change.
I will deal first with linear decompositions.

@@

There are three kinds of linear decompositions we have to deal with: zero,
addition, and scaling; corresponding to bunched connectives $I^*$, $\sep$, and
$\gr r \cdot {}$, respectively.
In each case, we have a simple preservation lemma, transforming an environment
of type $\Gamma \env\V \Delta$ and a decomposition of $\Delta$ into a
decomposition of $\Gamma$ and environments for all of the decomposed fragments
of $\Gamma$ and $\Delta$.

\begin{lemma}[environments preserve zero]\label{thm:lr-env-zero}
  Given an environment of type $\grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \gr 0$, we also have that $\grP \leq \gr 0$.
\end{lemma}
\begin{proof}
  $\grP \leq \grQ\gr\Psi \leq \gr0\gr\Psi = \gr0$, by environment
  compatibility and monotonicity and linearity of $\gr\Psi$.
\end{proof}

\begin{lemma}[environments preserve addition]\label{thm:lr-env-add}
  Given an environment of type $\grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \grQl + \grQr$ for some $\grQl$ and $\grQr$, we also have $\grPl$
  and $\grPr$ such that $\grP \leq \grPl + \grPr$ and there are environments
  of types $\grPl\gamma \env\V \grQl\delta$ and
  $\grPr\gamma \env\V \grQr\delta$.
\end{lemma}
\begin{proof}
  Let $\grPl \coloneqq \grQl\gr\Psi$ and $\grPr \coloneqq \grQr\gr\Psi$.
  Then, $\grP \leq \grQ\gr\Psi \leq \plr{\grQl + \grQr}\gr\Psi =
  \grQl\gr\Psi + \grQr\gr\Psi = \grPl + \grPr$, satisfying the first condition.
  Because clearly $\grPl \leq \grQl\gr\Psi$ and $\grPr \leq \grQr\gr\Psi$,
  \cref{thm:env-resize} on the original environment gives us the required
  pair of new environments.
\end{proof}

\begin{lemma}[environments preserve scaling]\label{thm:lr-env-scale}
  Given an environment of type $\grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \gr r\grQprime$ for some $\grQprime$, we also have a $\grPprime$
  such that $\grP \leq \gr r\grPprime$ and there is an environment of type
  $\grPprime\gamma \env\V \grQprime\delta$.
\end{lemma}
\begin{proof}
  Let $\grPprime \coloneqq \grQprime\gr\Psi$.
  Then, $\grP \leq \grQ\gr\Psi \leq \plr{\gr r\grQprime}\gr\Psi =
  \gr r\plr{\grQprime\gr\Psi} = \gr r\grPprime$, satisfying the first condition.
  Because clearly $\grPprime \leq \grQprime\gr\Psi$,
  \cref{thm:env-resize} on the original environment gives us the required
  new environment.
\end{proof}

Finally, I will also take the opportunity to give the bind lemma, allowing
environments to incorporate newly bound variables.
In the intuitionistic case, the bind lemma had two requirements on $\V$: $\V$
admits weakening and we can map variables into $\V$-values.
With usage annotations, the former is unreasonable, but it turns out that we
only need weakening by variables whose usage annotation is less than or equal
to $\gr0$.
The latter stays as-is, with the note that ``variable'' now means a
usage-checked variable.

\begin{lemma}[bind]\label{thm:lr-bind}
  Given functions
  ${\swarrow^k} : \forall \Gamma, \grR, \theta.~\grR \leq \gr0 \to
  \forallb{\V\,\Gamma \dotto \V\,\plr{\Gamma, \grR\theta}}$ and
  $\mathrm{vr} : \forallb{{\sqni} \dotto \V}$, we can turn an environment of
  type $\Gamma \env\V \Delta$ into an environment of type
  $\Gamma, \Theta \env\V \Delta, \Theta$ for any context $\Theta$.
\end{lemma}
\begin{proof}
  Let $\grP\gamma \coloneqq \Gamma$, $\grQ\delta \coloneqq \Delta$, and
  $\grR\theta \coloneqq \Theta$.
  Let the new linear map $\gr\Psi\gr' : \Ann^{\size\Delta + \size\Theta} \to
  \Ann^{\size\Gamma + \size\Theta}$ be $\gr\Psi \oplus \gr I$.
  That is, in block matrix notation,
  $\begin{pmatrix} \gr\Psi & \gr0 \\ \gr0 & \gr I \end{pmatrix}$.
  Checking that this linear map fits, we have
  $\begin{pmatrix}\grP & \grR\end{pmatrix}
  \leq \begin{pmatrix}\grQ\gr\Psi & \grR\gr I\end{pmatrix}
  = \begin{pmatrix}\grQ & \grR\end{pmatrix}\plr{\gr\Psi \oplus \gr I}$.
  For the action on variables, we are given vectors $\grPprime$,
  $\grR\gr'_\grP$, $\grQprime$, and $\grR\gr'_\grQ$ such that
  $\begin{pmatrix} \grPprime & \grR\gr'_\grP \end{pmatrix} \leq
  \begin{pmatrix} \grQprime & \grR\gr'_\grQ \end{pmatrix}
  \plr{\gr\Psi \oplus \gr I}$ and we have a variable of type
  $\grQprime\delta, \grR\gr'_\grQ\theta \sqni A$ for some type $A$.
  The constraint on the new vectors reduces to $\grPprime \leq \grQprime\gr\Psi$
  and $\grR\gr'_\grP \leq \grR\gr'_\grQ$.
  From the variable we either have a variable $x$ in $\delta$ with
  $\grQprime \leq \langle x \rvert$ and $\grR\gr'_\grQ \leq \gr0$, or a
  variable $y$ in $\theta$ with $\grQprime \leq \gr0$ and
  $\grR\gr'_\grQ \leq \langle y \rvert$.
  In the former case, the action of the original environment on $x$ gives us a
  $\V$-value in $\grPprime\gamma$, and the $\gr0$-weakening principle
  $\swarrow^k$, noting that $\grR\gr'_\grP \leq \grR\gr'_\grQ \leq \gr0$, gives
  us a $\V$-value in $\grPprime\gamma, \grR\gr'_\grP\theta$.
  In the latter case, we have that
  $\begin{pmatrix} \grPprime & \grR\gr'_\grP \end{pmatrix}
  \leq \begin{pmatrix} \grQprime\gr\Psi & \grR\gr'_\grQ \end{pmatrix}
  \leq \begin{pmatrix} \gr0\gr\Psi & \langle y \rvert \end{pmatrix}
  = \begin{pmatrix} \gr0 & \langle y \rvert \end{pmatrix}
  = \left\langle {\searrow}y \right\rvert$, so $y$ also serves as a
  usage-checked variable in $\grPprime\gamma, \grR\gr'_\grP\theta$.
  From this usage-checked variable, we get a $\V$-value in the same context
  using $\mathrm{vr}$.
\end{proof}

The requirements for identity and composition of environments look a bit like
the unit and lift of a Kleisli triple.
\todo{Change from equality to inequality}

\begin{lemma}[Identity environment]
  Given a function $\mathrm{vr} : \forallb{{\sqni} \dotto \V}$, for any
  $\Gamma$ we have an environment of type $\Gamma \env\V \Gamma$.
\end{lemma}
\begin{proof}
  Let $\gr\Psi$ be the identity map, which clearly satisfies
  $\grP = \grP\gr\Psi$.
  When acting on a variable, the equation $\grPprime = \grQprime\gr\Psi$ means
  that $\grPprime = \grQprime$, so we want, from a variable of type
  $\grPprime\gamma \sqni A$, a value of type $\V\,\grPprime\gamma\,A$, which
  we can get from $\mathrm{vr}$.
\end{proof}

\begin{lemma}\label{thm:env-comp-lemma}
  Given an environment $\rho : \Gamma \env\U \Delta$ for which we have, for any
  $\grPprime$ and $\grQprime$ such that
  $\grPprime = \grQprime\plr{\rho.\gr\Psi}$, we have a function
  $\mathrm{lift}_\rho :
  \forallb{\V\,\grQprime\delta \dotto \W\,\grPprime\gamma}$,
  we can map environments of type $\Delta \env\V \Theta$ into environments of
  type $\Gamma \env\W \Theta$.
\end{lemma}
\begin{proof}
  Let $\rho$ be as in the statement, and let $\sigma : \Delta \env\V \Theta$.
  For the environment we are constructing, let
  $\gr\Psi \coloneqq \sigma.\gr\Psi; \rho.\gr\Psi$, noting that
  $\grP = \grQ\plr{\rho.\gr\Psi} =
  \plr{\grR\plr{\sigma.\gr\Psi}}\plr{\rho.\gr\Psi}$.
  For the action on variables, we are given $\grPprime = \grRprime\gr\Psi$ with
  $\grRprime\theta \sqni A$.
  We can immediately apply the action of $\sigma$, giving us a value of type
  $\V\,\plr{\grRprime\plr{\sigma.\gr\Psi}}\,A$.
  We note that
  $\grPprime = \plr{\grRprime\plr{\sigma.\gr\Psi}}\plr{\rho.\gr\Psi}$, and
  apply $\mathrm{lift}_\rho$ to get the desired value.
\end{proof}

\begin{corollary}[Composition of environments]
  Given a function
  $\mathrm{lift} : \plr{\rho : \grP\gamma \env\U \grQ\delta} \to
  \forall \grPprime, \grQprime.~\grPprime = \grQprime\plr{\rho.\gr\Psi} \to
  \forallb{\V\,\grQprime\delta \dotto \W\,\grPprime\gamma}$, then we can
  compose environments of types $\Gamma \env\U \Delta$ and
  $\Delta \env\V \Theta$ into an environment of type $\Gamma \env\W \Theta$.
\end{corollary}

\todo{What do these choices mean?}
\begin{example}
  We can derive the following instances of environment composition.
  \begin{itemize}
    \item If $\U = \V = \W = {\sqni}$, then $\mathrm{lift}$ is given by the
      action of the renaming $\rho$ on variables.
      This allows us to derive composition of renamings.
    \item More generally, if $\V = {\sqni}$ and $\U = \W$, we can still use
      the action of the environment $\rho$.
      This means that renamings post-compose with any other sort of environment.
    \item If $\V = \W = {\vdash}$, then $\mathrm{lift}$ is given by a
      syntactic traversal.
      For example, if $\U = {\sqni}$, we need the action of renaming on terms
      to show that a renaming followed by a substitution composes to a
      substitution.
      If $\U = {\vdash}$, then the action of substitution on terms gives us that
      substitutions compose.
    \item More generally, if $\V = {\vdash}$ and we have a semantics from
      $\U$ to $\W$, then $\mathrm{lift}$ can be given by the semantic traversal
      of terms.
  \end{itemize}
\end{example}

>>>>

I settle on \cref{def:lr-env}, and prove various properties about it.

@@

We could, as in \cref{def:lr-rec-env}, use these three clauses to define what an
environment is.
However, I find them difficult to work with, as it is often easier to do
linear algebraic proofs separately from the rest of an environment.
For example, for identity and composition of environments, as we are about to
see, our working definition (\cref{def:lr-env})
is easier to use because we can rely on the closure of linear maps under
identity and composition.
Concretely, an inductive construction of, for example, an identity environment
would involve constructing an environment of type
$\grP\gamma, \grQ\delta \env\V \grP\gamma, \grQ\delta$ by constructing
environments of types $\grP\gamma, \gr0\delta \env\V \grP\gamma$ and
$\gr0\gamma, \grQ\delta \env\V \grQ\delta$.
These are not identity environments, so we would have to strengthen the
induction hypothesis.
Similar strengthening of the induction hypothesis would be required for
composition.

One of the primary test cases for environments is simultaneous substitution,
which will look like the \TirName{sub} rule below.
Note that we have taken $\V \coloneqq {\vdash}$ --- i.e.\ that the values
yielded by the environment are terms, namely the terms to be substituted in for
the free variables of the derivation of $\Delta \vdash A$.

@@

The admissibility of substitution will be by induction on the derivation of
$\Delta \vdash A$, so we will need to be able to adapt any environment we are
given to work with any possible context of new premises yielded by the rules of
\cref{fig:lr-bunched}.
In the simply typed case, the only change to the context we encountered was the
binding of new variables.
With usage annotations, we furthermore have linear decompositions of the
context, necessitating changes to the environment whenever usage annotations
change.

There are three kinds of linear decompositions we have to deal with: zero,
addition, and scaling; corresponding to bunched connectives $I^*$, $\sep$, and
$\gr r \cdot {}$, respectively.
In each of these three cases, we have a simple preservation lemma, transforming
an environment
of type $\Gamma \env\V \Delta$ and a decomposition of $\Delta$ into a
decomposition of $\Gamma$ and environments for all of the decomposed fragments
of $\Gamma$ and $\Delta$.

\begin{lemma}[environments preserve zero]\label{thm:lr-env-zero}
  Given an environment $\rho : \grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \gr 0$, we also have that $\grP \leq \gr 0$.
\end{lemma}
\begin{proof}
  $\grP \leq \grQ\gr\Psi \leq \gr0\gr\Psi = \gr0$, by environment
  compatibility from $\rho$ and monotonicity and linearity of $\gr\Psi$.
\end{proof}

\begin{lemma}[environments preserve addition]\label{thm:lr-env-add}
  Given an environment $\rho : \grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \grQl + \grQr$ for some $\grQl$ and $\grQr$, we also have $\grPl$
  and $\grPr$ such that $\grP \leq \grPl + \grPr$ and there are environments
  $\rho_l : \grPl\gamma \env\V \grQl\delta$ and
  $\rho_r : \grPr\gamma \env\V \grQr\delta$.
\end{lemma}
\begin{proof}
  Let $\grPl \coloneqq \grQl\gr\Psi$ and $\grPr \coloneqq \grQr\gr\Psi$.
  Then, $\grP \leq \grQ\gr\Psi \leq \plr{\grQl + \grQr}\gr\Psi =
  \grQl\gr\Psi + \grQr\gr\Psi = \grPl + \grPr$, satisfying the first condition.
  Because clearly $\grPl \leq \grQl\gr\Psi$ and $\grPr \leq \grQr\gr\Psi$,
  applying \cref{thm:env-resize} to $\rho$ gives us the required
  new environments $\rho_l$ and $\rho_r$.
\end{proof}

\begin{lemma}[environments preserve scaling]\label{thm:lr-env-scale}
  Given an environment $\rho : \grP\gamma \env\V \grQ\delta$ such that
  $\grQ \leq \gr r\grQprime$ for some $\grQprime$, we also have a $\grPprime$
  such that $\grP \leq \gr r\grPprime$ and there is an environment
  $\rho' : \grPprime\gamma \env\V \grQprime\delta$.
\end{lemma}
\begin{proof}
  Let $\grPprime \coloneqq \grQprime\gr\Psi$.
  Then, $\grP \leq \grQ\gr\Psi \leq \plr{\gr r\grQprime}\gr\Psi =
  \gr r\plr{\grQprime\gr\Psi} = \gr r\grPprime$, satisfying the first condition.
  Because clearly $\grPprime \leq \grQprime\gr\Psi$,
  applying \cref{thm:env-resize} to $\rho$ gives us the required
  new environment $\rho'$.
\end{proof}

The final change environments need to preserve is the binding of new free
variables.
In \cref{sec:syntactic-kits}, we had the operation \AgdaFunction{bindEnv} for
this purpose in the intuitionistic setting.
There, we relied on $\V$ supporting a map from $\ni$-variables and admitting
weakening.
In the usage-annotated setting, the former requirement is updated to having a
map from usage-checked $\sqni$-variables.
As for the latter requirement, it turns out that we only need $\V$ to admit
weakening by $\gr0$-annotated variables, which is much more reasonable than
general weakening.
\Cref{thm:lr-bind} adapts \AgdaFunction{bindEnv} for the usage-annotated
setting.

\begin{lemma}[\AgdaFunction{bindEnv}]\label{thm:lr-bind}
  Given functions
  ${\swarrow^k} : \forall \Gamma, \grR, \theta.~\grR \leq \gr0 \to
  \forallb{\V\,\Gamma \dotto \V\,\plr{\Gamma, \grR\theta}}$ and
  $\mathrm{vr} : \forallb{{\sqni} \dotto \V}$, we can turn an environment of
  type $\Gamma \env\V \Delta$ into an environment of type
  $\Gamma, \Theta \env\V \Delta, \Theta$ for any context $\Theta$.
\end{lemma}
\begin{proof}
  Let $\grP\gamma \coloneqq \Gamma$, $\grQ\delta \coloneqq \Delta$, and
  $\grR\theta \coloneqq \Theta$.
  Let the new linear map $\gr\Psi\gr' : \Ann^{\size\Delta + \size\Theta} \to
  \Ann^{\size\Gamma + \size\Theta}$ be $\gr\Psi \oplus \gr I$.
  That is, in block matrix notation,
  $\begin{pmatrix} \gr\Psi & \gr0 \\ \gr0 & \gr I \end{pmatrix}$.
  Checking that this linear map fits, we have
  $\begin{pmatrix}\grP & \grR\end{pmatrix}
  \leq \begin{pmatrix}\grQ\gr\Psi & \grR\gr I\end{pmatrix}
  = \begin{pmatrix}\grQ & \grR\end{pmatrix}\plr{\gr\Psi \oplus \gr I}$.
  For the action on variables, we are given vectors $\grPprime$,
  $\grR\gr'_\grP$, $\grQprime$, and $\grR\gr'_\grQ$ such that
  $\begin{pmatrix} \grPprime & \grR\gr'_\grP \end{pmatrix} \leq
  \begin{pmatrix} \grQprime & \grR\gr'_\grQ \end{pmatrix}
  \plr{\gr\Psi \oplus \gr I}$ and we have a variable of type
  $\grQprime\delta, \grR\gr'_\grQ\theta \sqni A$ for some type $A$.
  The constraint on the new vectors reduces to $\grPprime \leq \grQprime\gr\Psi$
  and $\grR\gr'_\grP \leq \grR\gr'_\grQ$.
  From the variable we either have a variable $x$ in $\delta$ with
  $\grQprime \leq \langle x \rvert$ and $\grR\gr'_\grQ \leq \gr0$, or a
  variable $y$ in $\theta$ with $\grQprime \leq \gr0$ and
  $\grR\gr'_\grQ \leq \langle y \rvert$.
  In the former case, the action of the original environment on $x$ gives us a
  $\V$-value in $\grPprime\gamma$, and the $\gr0$-weakening principle
  $\swarrow^k$, noting that $\grR\gr'_\grP \leq \grR\gr'_\grQ \leq \gr0$, gives
  us a $\V$-value in $\grPprime\gamma, \grR\gr'_\grP\theta$.
  In the latter case, we have that
  $\begin{pmatrix} \grPprime & \grR\gr'_\grP \end{pmatrix}
  \leq \begin{pmatrix} \grQprime\gr\Psi & \grR\gr'_\grQ \end{pmatrix}
  \leq \begin{pmatrix} \gr0\gr\Psi & \langle y \rvert \end{pmatrix}
  = \begin{pmatrix} \gr0 & \langle y \rvert \end{pmatrix}
  = \left\langle {\searrow}y \right\rvert$, so $y$ also serves as a
  usage-checked variable in $\grPprime\gamma, \grR\gr'_\grP\theta$.
  From this usage-checked variable, we get a $\V$-value in the same context
  using $\mathrm{vr}$.
\end{proof}

I put together the preceeding pieces to give a syntactic traversal operation over
$\name$ in the following section.
For the rest of this section, I observe some more constructions purely on
environments --- in particular, composition of environments given certain
assumptions about the families of values.

Following \citet{ACU15}, we expect (intuitionistic) ST$\lambda$C syntax to form
a relative monad over $\ni$ seen as a functor from the category of contexts
under renaming to the functor category $\blr{\mathrm{Ty}, \Set}$, where
$\mathrm{Ty}$ is the discrete category of ST$\lambda$C types.
Notice that, given $F,G : \blr{\mathrm{Ty}, \Set}$, a morphism from $F$ to $G$
is a function of type $\forallb{F \dotto G}$ (with naturality being trivial).
Therefore, we expect a relative monad, given as a Kleisli triple, to have a unit
$\eta_\Gamma : \forallb{\Gamma \ni \plr{-} \dotto \Gamma \vdash \plr{-}}$
given by the variable rule, and a Kleisli extension operator
${^*}_{\Gamma,\Delta} :
\forallb{\Gamma \ni \plr{-} \dotto \Delta \vdash \plr{-}} \to
\forallb{\Gamma \vdash \plr{-} \dotto \Delta \vdash \plr{-}}$
given by substitution.
Composition of substitutions falls out of this framework as Kleisli composition.
However, in the usage-aware case, substitution needs not just a mapping of
variables $f : \forallb{\Gamma \sqni \plr{-} \dotto \Delta \vdash \plr{-}}$,
but rather an environment $\rho : \Delta \env\vdash \Gamma$, as we have already
discussed.
It therefore makes sense for our replacement for the Kleisli extension operator
to similarly take an environment rather than a simple variable mapping.

\Cref{thm:env-comp-lemma} below amounts to deriving a modified notion of Kleisli
composition from a modified Kleisli extension.
Additionally, \cref{thm:env-id-lemma} is required to turn a monadic unit into an
identity environment.
Both lemmas are stated in terms of general $\U$/$\V$/$\W$-environments, with
some specific examples (e.g.\ for renaming and substitution) below them.

\begin{lemma}[Identity environment]\label{thm:env-id-lemma}
  Given a function
  \[
    \mathrm{vr}_{\Gamma'} : \forallb{\Gamma' \sqni \plr{-} \dotto \V\,\Gamma'}
  \]
  for any
  $\Gamma$ we have an environment $\mathrm{id} : \Gamma \env\V \Gamma$.
\end{lemma}
\begin{proof}
  Let $\Gamma = \grP\gamma$.
  Let $\gr\Psi$ be the identity map, which clearly satisfies
  $\grP \leq \grP\gr\Psi$.
  When acting on a variable, the inequality $\grPprime \leq \grQprime\gr\Psi$
  means that $\grPprime \leq \grQprime$.
  We are given a variable of type $\grQprime\gamma \sqni A$, which we can
  coerce to a variable of type $\grPprime\gamma \sqni A$, upon which we apply
  $\mathrm{vr}$ to get the required value of type $\V\,\grPprime\gamma\,A$.
\end{proof}

\begin{lemma}[Composition of environments]\label{thm:env-comp-lemma}
  Given a function
  \[
    \mathrm{lift}_{\Gamma', \Delta'} :
    \Gamma' \env\U \Delta' \to \forallb{\V\,\Delta' \dotto \W\,\Gamma'}
  \]
  we can compose environments $\rho : \Gamma \env\U \Delta$ and
  $\sigma : \Delta \env\V \Theta$ into an environment
  $\rho \gg \sigma : \Gamma \env\W \Theta$.
\end{lemma}
\begin{proof}
  Let $\Gamma = \grP\gamma$, $\Delta = \grQ\delta$, and $\Theta = \grR\theta$.
  Take $\gr\Psi$ to be the composition $\plr{\sigma.\gr\Psi}\plr{\rho.\gr\Psi}$,
  noting that
  $\grP \leq \grQ\plr{\rho.\gr\Psi} \leq
  (\grR\plr{\sigma.\gr\Psi})\plr{\rho.\gr\Psi} = \grR\gr\Psi$
  thanks to the inequalities yielded by $\sigma$ and $\rho$.
  When acting on a variable, we are given $\grPprime \leq \grRprime\gr\Psi$ and
  a variable $v : \grRprime\theta \sqni A$, and want a value of type
  $\W\,\grPprime\gamma\,A$.
  Let $\grQprime \coloneqq \grRprime\plr{\sigma.\gr\Psi}$, with inequality
  $\grQprime \leq \grRprime\plr{\sigma.\gr\Psi}$ giving us a value
  $\sigma(v) : \V\,\grQprime\delta\,A$.
  We wish to apply $\mathrm{lift}$ to $\sigma(v)$ with
  $\Gamma' \coloneqq \grPprime\gamma$ and $\Delta' \coloneqq \grQprime\delta$ to
  complete the construction of the $\W$-value.
  To do this, we need an environment of type
  $\grPprime\gamma \env\U \grQprime\delta$, which we can get from $\rho$ using
  \cref{thm:env-resize}, noting that
  $\grPprime \leq \grRprime\plr{\sigma.\gr\Psi}\plr{\rho.\gr\Psi} =
  \grQprime\plr{\rho.\gr\Psi}$.
\end{proof}

\begin{example}
  We can derive the following instances of environment composition.
  \begin{description}
    \item[Composition of renamings.]
      If $\U = \V = \W = {\sqni}$, then we can take
      $\mathrm{lift}\,\rho\,x \coloneqq \rho(x)$.
      This allows us to compose two renamings into a single renaming.
    \item[Post-composition with a renaming.]
      More generally, if $\V = {\sqni}$ and $\U = \W$, we can still use
      the action of $\rho$ as $\mathrm{lift}$.
      This means that we can do a composition
      $\Gamma \env\U \Delta \env\sqni \Theta$ to get an environment of type
      $\Gamma \env\U \Theta$.
    \item[Composition of substitutions.]
      If $\V = \W = {\vdash}$, then $\mathrm{lift}$ is given by a
      syntactic traversal (in the sense of \cref{sec:syntactic-kits} and
      \cref{sec:lrsub}).
      For example, if $\U = {\sqni}$, we need the action of renaming on terms
      to show that a renaming followed by a substitution composes to a
      substitution.
      If $\U = {\vdash}$, then the action of substitution on terms gives us that
      substitutions compose.
    \item[Composing semantics with substitution.]
      More generally, if $\V = {\vdash}$ and we have a semantics (in the sense of
      \cref{sec:gen-sem} and \cref{sec:framework}) from
      $\U$ to $\W$, then $\mathrm{lift}$ can be given by the semantic traversal
      of terms.
  \end{description}
\end{example}
